---
title: "Fulmore_v2"
author: "Eric V"
date: "`r Sys.Date()`"
output: html_document
---

---
title: "Fulmore River Sockeye Salmon Smolt Enumeration"
author: "Eric V"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Project Background
 
This is a live document intended to present preliminary information about juvenile salmon captured in the Fulmore River by rotary screw trap. The Fulmore River is located in the traditional territory of the Tlowitsis First Nation, at the head of Port Neville, approximately 80 NE of Campbell River. The trap was installed on April 11-12 and will fish roughly 5-days a week (Monday to Friday) until June 15, 2024. All fish captured in the trap will be identified to species and measured for fork length. Biological samples will be collected from a subset of fish each day. Releases of marked fish will occur weekly to assess trap efficiency. 

```{r prepare working environment, echo=FALSE, warning = FALSE, message = FALSE}
## Load Packages
    library(tidyverse)
    library(tidyr)
    library(dplyr)
    library(zoo)
    library(lubridate)
    library(reshape2)
    library(ggplot2)
    library(readxl)
    library(kableExtra)
    library(markdown)
    library(captioner)

## Define Working Directory ---
setwd("C:/Users/evogt/R Analysis/EAV/GitHubMarkdown/Fulmore/docs/")

## Define and Load Input Files
  #1. PacFish Temp and Flow Data
    #catch_dat <- "Data/RST Catch/RST_Catch_2024.05.02.xlsx"
    #catch_dat <- read_excel(catch_dat)

    catch_dat <- "Data/RST Catch/StandardFishCollectionComplete_29389-1.xlsx"
    data <- read_excel(catch_dat, sheet = "Fish Data")
 
  ## Define Output Locations
    plot.output <- "Plots/"
    model.output <- "Output/"

## Load Ecofish Plot Theme ----          
    plottheme = function() {
      theme(
        legend.key.size = unit(0.75, "lines"),
        legend.key = element_blank(),
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(size = 8, face = "bold"),
        axis.line = element_line(colour = "grey", size = 1),
        panel.grid.major.y = element_line(colour = "gray", linetype = "dashed"),
        panel.grid.major.x = element_line(colour = "gray"),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "gray40", fill = NA),
        panel.background = element_rect(fill = NA, colour = "gray40"),
        axis.title = element_text(face = "bold", size = 9),
        axis.text = element_text(size = 8, colour = "black"),
        axis.ticks.x = element_blank(),
        plot.margin = unit(c(2, 5, 2, 2), "mm"),
        strip.background = element_blank(),
        strip.text = element_text(size = 8, face = "bold")
      )
    }          

```


Since April 15, the trap has fished for a total of `r (length(unique(catch_dat$Date)))` days, resulting in the capture of Sockeye Salmon, Coho Salmon, Chum Salmon, Coastal Cutthroat Trout, Dolly Varden Char, Pacific Lamprey and Sculpin (Cottidae spp.).

 
```{r tallytable, echo = FALSE, message = FALSE, warning = FALSE}
## Prepare Data
    catch_dat <- data %>%                   
                 rename(Length = "Measured Length (mm)",
                        Weight = "Weight (g)",
                        Date = "Date (YYYY-MM-DD)") %>%
                 mutate(Condition = 100*(Weight/(Length^3)),
                        Species = recode(as.factor(Species), 
                                             SK        = "Sockeye Salmon",
                                             CO        = "Coho Salmon",
                                             CM        = "Chum Salmon",
                                             CT        = "Coastal Cutthroat",
                                             'BT/DV'   = "Dolly Varden",
                                             CC        = "Coastal Sculpin",
                                             L   = "Pacific Lamprey"),
                             Species = factor(Species, 
                                              levels = c("Sockeye Salmon", "Coho Salmon", "Chum Salmon", 
                                                         "Coastal Cutthroat", "Dolly Varden", "Coastal Sculpin", 
                                                         "Pacific Lamprey")))
    
    tally.table <- catch_dat %>% 
                      group_by(Species) %>%
                              summarize(count = n(),
                                        nFL     = sum(Length >0, na.rm = TRUE),
                                        meanFL  = mean(Length, na.rm = TRUE),
                                        minFL   = min(Length, na.rm = TRUE),
                                        maxFL   = max(Length, na.rm = TRUE),
                                        nWT     = sum(Weight > 0, na.rm=TRUE),
                                        meanWT  = mean(Weight, na.rm = TRUE),
                                        minWT   = min(Weight, na.rm = TRUE),
                                        maxWT   = max(Weight, na.rm = TRUE))

## Prepare Kable Table
    kable(tally.table,
          "html",
          booktabs = T,
          align = c("c"),
          col.names = c("Species", "Total \n Captures","n", "Mean", "Min", "Max", "n", "Mean", "Min", "Max"),
          caption = "**Summary of total number and biological data collected of fish captured since April 15, 2024.",
          digits = 1) %>%
      #column_spec(c(1, 3,10,17), bold = TRUE) %>%
      kable_styling("striped", full_width =  FALSE,
                    position = "center", 
                    font_size = 12,
                    fixed_thead = TRUE) %>%
      add_header_above(c(" " = 2, "Fork Length" = 4, "Weight (g)" = 4)) 
```

## Length Frequency Plots

```{r length frequency plot for all coho and sockeye, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Length-frequency distribution for Sockeye Salmon captured in the Fulmore River.", message = FALSE, fig.dim=c(10,18)}
## Total tally by fork length
all_catches <- catch_dat %>% 
               filter(Species %in% c("Sockeye Salmon" ,"Coho Salmon")) %>%
               group_by(Species, Length) %>%
                   summarize(Count = n()) %>%
                   mutate(Date = "All Dates") %>%
                   select("Species","Length","Count") 
               

all_catch_LFreq.plot <- ggplot(all_catches, aes(x = Length, y = Count))+
                        geom_col() +
                        labs(x = "Fork Length (mm)", y = "Frequency") + 
                        scale_x_continuous(breaks = seq(20, 210, by = 20))+
                        facet_grid(Species~.) +
                        plottheme()

all_catch_LFreq.plot
```


## Sockeye Salmon

The length-frequency distribution for Sockeye Salmon smolts is multimodal (separating at ~120 mm), indicating at least two groups of different sized fish are emigrating simultaneously. Scale aging will be required to assess what may be contributing to this observed size difference. However, it is likely best explained by a difference in age class between modes (e.g., Age-1, Age-2 and Age-3 smolts), though it is also possible that some of the difference is explained by a difference in productivity between rearing environments (e.g., some smolts may be the same age, but larger smolts reared in more productive environments than smaller smolts). 

A plot of length vs weight may help explain what is going on. If this plot showed two distinct clusters it would be reasonable to assume that the difference in length was due primarily to differences in age between smolts. 

```{r sockeye scatter plot, echo = FALSE, message = FALSE, warning = FALSE}
SK_dat <- catch_dat %>% 
              filter(Species == "Sockeye Salmon") %>%
              mutate(AgeClass =ifelse(Length>=175,"Age-3",
                                      ifelse(Length>120,"Age-2","Age-1")),
                     TotalCount = n())

ggplot(SK_dat) + 
      geom_point(aes(x = Length, y = Weight, color = AgeClass)) +
      plottheme()


```


Well that is very interesting. The scatter plot suggests that there may be three distinct age classes AND that fish may be rearing in two unique environments. 

```{r}
ggplot(SK_dat) + 
  geom_point(aes(x = Length, y = Condition, color = AgeClass)) +
  plottheme()
```

