---
title: "Fulmore River Sockeye Salmon Smolt Enumeration"
author: "Eric V"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Project Background
 
This is a live document intended to present preliminary information about juvenile salmon captured in the Fulmore River by rotary screw trap. The Fulmore River is located in the traditional territory of the Tlowitsis First Nation, at the head of Port Neville, approximately 80 NE of Campbell River. The trap was installed on April 11-12, 2024 and fished ~5-days a week (Monday to Friday) until May 31, 2024. All fish captured in the trap were identified to species and measured for fork length. Biological samples will be collected from a subset of fish each day. Releases of marked fish occurred weekly to assess trap efficiency. No estimate of capture efficiency is presented as smolts did not take to Bismark (would not dye) and captures were too low to effectively mark with fin clip. 

```{r prepare working environment, echo=FALSE, warning = FALSE, message = FALSE}
## Load Packages
    library(tidyverse)
    library(tidyr)
    library(dplyr)
    library(zoo)
    library(lubridate)
    library(reshape2)
    library(ggplot2)
    library(readxl)
    library(kableExtra)
    library(markdown)
    library(captioner)
    fig_nums <- captioner()

## Define Working Directory ---
setwd("C:/Users/evogt/R Analysis/EAV/GitHubMarkdown/Fulmore/docs/")

## Define and Load Input Files
  #1. PacFish Temp and Flow Data
    #catch_dat <- "Data/RST Catch/RST_Catch_2024.05.02.xlsx"
    #catch_dat <- read_excel(catch_dat)

    catch_dat <- "Data/RST Catch/StandardFishCollectionComplete_29389-1.xlsx"
    data <- read_excel(catch_dat, sheet = "Fish Data")
 
  ## Define Output Locations
    plot.output <- "Plots/"
    model.output <- "Output/"

## Load Ecofish Plot Theme ----          
    plottheme = function() {
      theme(
        legend.key.size = unit(0.75, "lines"),
        legend.key = element_blank(),
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(size = 8, face = "bold"),
        axis.line = element_line(colour = "grey", size = 1),
        panel.grid.major.y = element_line(colour = "gray", linetype = "dashed"),
        panel.grid.major.x = element_line(colour = "gray"),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "gray40", fill = NA),
        panel.background = element_rect(fill = NA, colour = "gray40"),
        axis.title = element_text(face = "bold", size = 9),
        axis.text = element_text(size = 8, colour = "black"),
        axis.ticks.x = element_blank(),
        plot.margin = unit(c(2, 5, 2, 2), "mm"),
        strip.background = element_blank(),
        strip.text = element_text(size = 8, face = "bold")
      )
    }          

```


```{r prep catch_dat,, echo = FALSE, message = FALSE, warning = FALSE}
## Prepare Data
catch_dat <- data %>%                   
                 rename(Length = "Measured Length (mm)",
                        Length.Est = "Estimated Length (mm)",
                        Weight = "Weight (g)",
                        Date = "Date (YYYY-MM-DD)") %>%
                 mutate(Condition = 100*(Weight/(Length^3)),
                        Species = recode(as.factor(Species), 
                                             SK        = "Sockeye Salmon",
                                             CO        = "Coho Salmon",
                                             CM        = "Chum Salmon",
                                             CT        = "Coastal Cutthroat",
                                             'BT/DV'   = "Dolly Varden",
                                             CC        = "Coastal Sculpin",
                                             L         = "Pacific Lamprey",
                                            TSB        = "Three-Spine Stickleback"),
                             Species = factor(Species, 
                                              levels = c("Sockeye Salmon", "Coho Salmon", "Chum Salmon", 
                                                         "Coastal Cutthroat", "Dolly Varden", "Coastal Sculpin", 
                                                         "Pacific Lamprey", "Three-Spine Stickleback")))
    
spp_summary <- catch_dat %>% 
                  group_by(Species) %>%
                  summarize(Count = n())

```

## Captures to Date

Since April 15, the trap has fished for a total of `r (length(unique(catch_dat$Date)))` days, resulting in the capture of Sockeye Salmon, Coho Salmon, one Chum Salmon, Coastal Cutthroat Trout, Dolly Varden Char, Pacific Lamprey, Sculpin (Cottidae spp.) and sticklebacks.
 
```{r tallytable, echo = FALSE, message = FALSE, warning = FALSE}
## Prepare summary table
    tally.table <- catch_dat %>% 
                      group_by(Species) %>%
                              summarize(count = n(),
                                        nFL     = sum(Length >0, na.rm = TRUE),
                                        meanFL  = mean(Length, na.rm = TRUE),
                                        minFL   = min(Length, na.rm = TRUE),
                                        maxFL   = max(Length, na.rm = TRUE),
                                        nWT     = sum(Weight > 0, na.rm=TRUE),
                                        meanWT  = mean(Weight, na.rm = TRUE),
                                        minWT   = min(Weight, na.rm = TRUE),
                                        maxWT   = max(Weight, na.rm = TRUE))

## Prepare Kable Table
    kable(tally.table,
          "html",
          booktabs = T,
          align = c("c"),
          col.names = c("Species", "Total \n Captures","n", "Mean", "Min", "Max", "n", "Mean", "Min", "Max"),
          caption = "Summary of total number and biological data collected of fish captured since April 15, 2024.",
          digits = 1) %>%
      #column_spec(c(1, 3,10,17), bold = TRUE) %>%
      kable_styling("striped", full_width =  FALSE,
                    position = "center", 
                    font_size = 12,
                    fixed_thead = TRUE) %>%
      add_header_above(c(" " = 2, "Fork Length" = 4, "Weight (g)" = 4)) 
```

# Sockeye Salmon

The length-frequency distribution for Sockeye Salmon smolts is multimodal (separating at ~120 mm), indicating at least two groups of different sized fish are emigrating simultaneously. Scale aging will be required to verify whether this difference is due to different age classes, though it is also possible that some of the variability is explained by more than just age, such as a difference in rearing environments, life history strategies, or even sampling error! 

```{r length frequency plot for all sockeye, echo = FALSE, message = FALSE, warning = FALSE, fig.dim=c(10,8)}
## Total tally by fork length
SK_LF_data <- catch_dat %>% 
#all_catches <- catch_dat %>%   
               filter(Species == "Sockeye Salmon") %>%
               group_by(Species, Length) %>%
                   summarize(Count = n()) %>%
                   mutate(Date = "All Dates") %>%
                   select("Species","Length","Count") 
               

all_catch_LFreq.plot <- ggplot(SK_LF_data, aes(x = Length, y = Count))+
                        geom_col() +
                        labs(x = "Fork Length (mm)", y = "Frequency") + 
                        scale_x_continuous(breaks = seq(20, 210, by = 5),
                                           minor_breaks = seq(20, 210, by = 2))+
                        facet_grid(Species~.) +
                        plottheme()

all_catch_LFreq.plot
```
`r fig_nums("all_catch_LFreq.plot", "Length frequency plot of Sockeye Salmon captured in the Fulmore River.")`

```{r smolt proportion, echo = FALSE, warning = FALSE, message = FALSE}
SK_dat <- catch_dat %>% 
              filter(Species == "Sockeye Salmon") %>%
              mutate(AgeClass =ifelse(Length>=175,"Age-3",
                                      ifelse(Length>120,"Age-2","Age-1")),
                     TotalCount = n()) %>%
              filter(!is.na(AgeClass))

SK_prop <- SK_dat %>%
              group_by(AgeClass,TotalCount) %>%
              summarize(Count  = n(),
                        nFL = sum(!is.na(Length)),
                        MeanFL = mean(Length, na.rm = TRUE),
                        SEFL   = sd(Length, na.rm = TRUE)/sqrt(Count),
                        MinFL  = min(Length, na.rm = TRUE),
                        MaxFL  = max(Length, na.rm = TRUE),
                        nWT    = sum(!is.na(Weight)),
                        MeanWT = mean(Weight, na.rm = TRUE),
                        SEWT   = sd(Weight, na.rm = TRUE)/sqrt(Count),
                        MinWT  = min(Weight, na.rm = TRUE),
                        MaxWT  = max(Weight, na.rm = TRUE))%>%
              mutate(Prop = (Count/TotalCount)*100) %>% 
              select("AgeClass" ,"Count","Prop","nFL", "MeanFL", "SEFL", "MinFL", "MaxFL", "nWT", "MeanWT","SEWT", "MinWT","MaxWT") %>%
              filter(AgeClass != "NA")


kable(SK_prop,
      "html",
      booktabs = T,
      align = c("c"),
      col.names = c("Age Class", "n", "Prop (%)","n" ,"Mean", "SE","Min", "Max", "n","Mean", "SE","Min", "Max"),
      caption = "**Estimated Size of Age-1, Age-2 and Age-3 Sockeye Salmon captured in the Fulmore River in spring 2024.**",
      digits = 1) %>%
  #column_spec(c(1, 3,10,17), bold = TRUE) %>%
  kable_styling("striped", full_width =  FALSE,
                position = "center", 
                font_size = 12,
                fixed_thead = TRUE) %>%
  add_header_above(c(" " = 3, "Fork Length" = 5, "Weight (g)" = 5)) 

```

### Data Exploration

To assess why the length-frequency plot has a multimodal distribution we can plot length vs weight. If this plot showed two distinct clusters it would be reasonable to assume that the difference in length was due primarily to differences in age between smolts (e.g., there would be a clear difference in the size of Age-1 and Age-2 fish). But interestingly, this is not what the length vs weight plot shows. Instead, the plot shows distinct differences in weight for each length, suggesting that there are likely two age-classes AND that smolts from both age classes are of different condition - one group that appears is better condition (where fish are heavier at a given length), and one group that is relatively poorer condition (where fish are lighter at a given length). The reason for this difference in condition is unknown.   

```{r sockeye scatter plot, echo = FALSE, message = FALSE, warning = FALSE}


SK_scatter <- ggplot(SK_dat) + 
                  geom_point(aes(x = Length, y = Weight, color = AgeClass)) +
                  plottheme()
SK_scatter

```

`r fig_nums("SK_scatter", "Fork length and weight of Sockeye Salmon captured in the Fulmore River.")`

This difference between groups becomes more apparent when we plot Fulton's condition factor (k) vs length. This figure (below) shows four distinct groups, and there looks to be a clear separation in condition factor at ~ 0.00115. Though very much a speculation, this break may be due to a difference in environmental conditions between rearing environments. 

```{r SK_ConditionScatter, echo=FALSE, message=FALSE, warning=FALSE}
sk_cond <- ggplot(SK_dat) + 
              geom_point(aes(x = Length, y = Condition, color = AgeClass)) +
              labs(x = "Fork Length (mm)", y = "Condition Factor (k)") +
              plottheme()
sk_cond

```

`r fig_nums("sk_cond", "Fork length and condition factor (k) of Sockeye Salmon captured in the Fulmore River.")`

```{r}
sk_condDate <- ggplot(SK_dat) + 
              geom_point(aes(x = Date, y = Condition, color = AgeClass)) +
              labs(x = "", y = "Condition Factor (k)") +
              plottheme()
sk_condDate
```

`r fig_nums("sk_condDate", "Condition Factor (k) of Sockeye Salmon over time in the Fulmore River.")`


If we apply this information on condition factor to our original length vs weight plot we can see that our initial age assignment (Age 1 < 120 mm, Age 2 > 120 mm) may also need to take groupings into account. Specifically, poorer condition fish in Group B appear to be a smaller at age than better condition fish in Group B.

```{r SK scatter w. condition and rearing env,  echo = FALSE, message = FALSE, warning = FALSE}
SK_dat2 <- SK_dat %>%
              mutate(RearingEnv = ifelse(Condition>=0.00115,"Group A", "Group B"),
                     RearingEnv = factor(RearingEnv, 
                                       levels = c("Group A", "Group B"))) %>%
              filter(!is.na(RearingEnv))


SK_RearEnv<- ggplot(SK_dat2) + 
                geom_point(aes(x = Length, y = Weight, color = AgeClass, shape = RearingEnv)) +
                scale_shape_manual(values = c(1, 3)) +
                labs(x = "Fork Length (mm)", y = "Weight (g)") +
                plottheme()

SK_RearEnv
```

`r fig_nums("SK_RearEnv", "Fork length and weight of Sockeye Salmon captured in the Fulmore River, with emphasis on condition factor.")`

The data suggests there is a physiological difference among smolts in terms of size and condition at age. Though not definitive, differences in behaviour would also be an indication that that there may be environmental or behavioural factors influencing observed differences in condition. To see if there is evidence for a behavioural difference between groups we can re-plot the length-frequency figure to show the number of fish from each group that were captured per week of trapping and look at the total number of fish from each group originating from each group. 


```{r SK_FL, echo = FALSE, warning = FALSE, message = FALSE, fig.dim=c(10,18)}
#, out.height="1200px"
## Prepare Plots ----
## * Length Frequency Plots  ----  
## ** Prepare Data
    ## Total tally by fork length
    SK_LF_data <- SK_dat2 %>% 
                          group_by(Length) %>%
                          summarize(Count = n()) %>%
                          mutate(Date = "All Dates") %>%
                          select("Date","Length","Count")

    ## Weekly tally by fork length
    SK_weekly_catch <- SK_dat2 %>%
                      mutate(Date = isoweek(ymd(Date)), # convert date to weeknumber
                             Date = recode(as.factor(Date),
                                           '16' = "April 15-21",
                                           '17' = "April 22-28",
                                           '18' = "April 29-May 5",
                                           '19' = "May 6-12"),
                             Date = factor(Date, 
                                           levels = c("April 15-21",
                                                      "April 22-28",
                                                      "April 29-May 5",
                                                      "May 6-12"))) %>%  
                      group_by(Date, RearingEnv, Length) %>%
                        summarize(Count = n()) %>%
                        select("Date","Length","Count") 


## ** Prepare Daily Length Frequency Plots
    ## All Catches
    all_fl.plot <- ggplot(SK_LF_data) +
                         geom_col(aes(x = Length, y = Count)) + 
                         labs(x = "Fork Length (mm)", y = "Frequency (# Fish)") +
                         scale_x_continuous(breaks = seq(70, 210, by = 2))+
                         scale_y_continuous(breaks = seq(0, 30, 2))+
                         plottheme() +
                         theme(axis.text.x = element_text(angle = 45, vjust = 1.2, hjust = 1)) 
 
    ## Faceted by Week
    SK_weekly_fl.plot <- ggplot(SK_weekly_catch) + 
                      geom_col(aes(x = Length, y = Count)) + 
                      labs(x = "Fork Length (mm)", y = "Frequency (# Fish)") +
                      scale_x_continuous(breaks = seq(70, 210, by = 10))+
                      scale_y_continuous(breaks = seq(0, 50, 2))+
                      plottheme() +
                      facet_grid(Date~RearingEnv)
    
## ** Select Figures to Display
    #all_fl.plot
    SK_weekly_fl.plot
``` 

`r fig_nums("SK_weekly_fl.plot", "Length-Frequency plot showing number of Sockeye Salmon captured per week from Group A (better condition) and Group B (poorer condition).")`


```{r plot of cumulative catches over time, echo = FALSE, warning = FALSE, message = FALSE,}
    cum.sk <- SK_dat2 %>%
              group_by(Date, RearingEnv) %>%
              summarize(Count = n(),
                        Prop  = Count/sum(Count),
                        cumsum = cumsum(Count)) %>%
              select("Date","RearingEnv","Count", "Prop", "cumsum") %>%
              mutate(Date = as.character(Date))

      cum.sk.plot <- ggplot(cum.sk) +
                       geom_col(aes(x = Date, y = cumsum)) + 
                       labs(x = "", y = "Frequency (# Fish)") +
                       #scale_x_continuous(breaks = seq(70, 210, by = 2))+
                       #scale_y_continuous(breaks = seq(0, 250, 25))+
                       plottheme() +
                       facet_grid(RearingEnv~.)
      
      cum.sk.plot
```

`r fig_nums("cum.sk.plot", "Number of Sockeye Salmon captured per day from Group A and Group B.")`

Well look at that! Though our data is incomplete and we clearly missed the start of the migration period, the preliminary results suggest that there may be a difference in migration timing between fish from each group; specifically, better condition fish (Group A) appear to migrate earlier than poorer condition fish from Group B.


### Sockeye Salmon Data Summary
Based on the above plots, it is reasonable to speculate that:
  1. There are at least 2 age classes of Sockeye Salmon smolts emigrating from the Fulmore watershed.
  2. Analysis of scale samples will be required to verify the age-bins assigned in this preliminary assessment.
  3. Sockeye smolts can be grouped by condition into two genearl groups: 1) better condition fish that appear to migrate earlier, and 2) relatively poorer condition fish that migrate later.  
  4. The bulk of the migration occurred prior to trap installation, so it is unclear whether current results are representative of the entire run.


# Coho Salmon

The length-frequency plot for Coho Salmon also shows two distinct age classes; Coho Salmon fry were ~30-65 mm while parr were ~85 to 145 mm. Coho Salmon fry may have been moving to the estuary or to more downstream freshwater rearing habitats, while parr are almost certainly migrating towards the sea. Similar to the length weight plot for Sockeye there appears to be a subset of Coho Salmon that are relatively poorer condition, though the majority appear to be of similar condition.    

```{r CO_FL, echo = FALSE, warning = FALSE, message = FALSE, out.width= "1500px"}

## Prepare Plots ----
## * Length Frequency Plots  ----  
## ** Prepare Data
    ## Total tally by fork length
    co.dat <- catch_dat %>% 
                      filter(Species == "Coho Salmon")


    co.all_catches <- co.dat %>% 
                      group_by(Length) %>%
                      summarize(Count = n()) %>%
                      mutate(Date = "All Dates") %>%
                      select("Date","Length","Count")
    
    
## ** Prepare Daily Length Frequency Plots
    ## Faceted by Day
    co.fl.plot <- ggplot(co.all_catches) + 
                         geom_col(aes(x = Length, y = Count)) + 
                         labs(x = "Fork Length (mm)", y = "Frequency (# Fish)") +
                         scale_x_continuous(breaks = seq(0, 150, by = 5))+
                         # scale_y_continuous(breaks = seq(0, 50, 2))+
                         plottheme() +
                         theme(axis.text.x = element_text(angle = 45, vjust = 1.2, hjust = 1)) +
                         facet_grid(Date ~.)

    ## ** Select Figures to Display
    
    co.fl.plot

```  

`r fig_nums("co.fl.plot", "Length-frequency distribution for all Coho Salmon captured in the Fulmore River.")`


```{r CO_Length_weight, echo = FALSE, warning = FALSE}
co.plot_flwt <- ggplot(co.dat) +
                geom_point(aes(x = Length, y = Weight)) +
                labs (x = "Fork Length (mm)", y = "Weight (g)") +
                plottheme()

co.plot_flwt
```

`r fig_nums("co.plot_flwt", "Length and weight of Coho Salmon captured in the Fulmore River.")`

