---
title: "Fulmore River Sockeye Salmon Smolt Enumeration"
author: "Eric V"
date: "`r Sys.Date()`"
output: 
  
  html_document:
    fig_caption: true
    code_folding: "show"
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Project Background

This is a live document intended to share information about juvenile salmon captured in the Fulmore River by rotary screw trap. The Fulmore River is located in the traditional territory of the Tlowitsis First Nation, at the head of Port Neville, approximately 80 NE of Campbell River. The trap was installed on April 11-12 and will fish roughly 5-days a week (Monday to Friday) until June 15, 2024. All fish captured in the trap will be identified to species and measured for fork length. Biological samples will be collected from a subset of fish each day. Releases of marked fish will occur weekly to assess trap efficiency. 

```{r prepare working environment, echo=FALSE, warning=FALSE, message = FALSE}
## Load Packages
    library(tidyverse)
    library(tidyr)
    library(dplyr)
    library(zoo)
    library(lubridate)
    library(reshape2)
    library(ggplot2)
    library(readxl)
    library(kableExtra)
    library(markdown)
    library(captioner)

## Define Working Directory ---
setwd("C:/Users/evogt/R Analysis/EAV/GitHubMarkdown/Fulmore/docs/")

## Define and Load Input Files
  #1. PacFish Temp and Flow Data
    catch_dat <- "Data/RST Catch/RST_Catch_2024.04.17.xlsx"
    catch_dat <- read_excel(catch_dat)
    
  ## Define Output Locations
    plot.output <- "Plots/"
    model.output <- "Output/"

## Load Ecofish Plot Theme ----          
    plottheme = function() {
      theme(
        legend.key.size = unit(0.75, "lines"),
        legend.key = element_blank(),
        legend.title = element_text(size = 9, face = "bold"),
        legend.text = element_text(size = 9),
        plot.title = element_text(size = 8, face = "bold"),
        axis.line = element_line(colour = "grey", size = 1),
        panel.grid.major.y = element_line(colour = "gray", linetype = "dashed"),
        panel.grid.major.x = element_line(colour = "gray"),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "gray40", fill = NA),
        panel.background = element_rect(fill = NA, colour = "gray40"),
        axis.title = element_text(face = "bold", size = 9),
        axis.text = element_text(size = 8, colour = "black"),
        axis.ticks.x = element_blank(),
        plot.margin = unit(c(2, 5, 2, 2), "mm"),
        strip.background = element_blank(),
        strip.text = element_text(size = 8, face = "bold")
      )
    }          

```

# Captures to Date

Since April 15, the trap has fished for a total of `r (length(unique(catch_dat$Date)))` days, resulting in the capture of Sockeye Salmon, Coho Salmon, Chum Salmon, Coastal Cutthroat Trout, Dolly Varden Char, Pacific Lamprey and Sculpin (Cottidae spp.)
 
```{r tallytable, echo = FALSE, message = FALSE}
## Prepare Data
tally.table <- catch_dat %>% 
                  mutate(Species = recode(as.factor(Species), 
                                         SK        = "Sockeye Salmon",
                                         CO        = "Coho Salmon",
                                         CM        = "Chum Salmon",
                                         CT        = "Coastal Cutthroat",
                                         'DV-BT'   = "Dolly Varden",
                                         CC        = "Coastal Sculpin",
                                         Lamprey   = "Pacific Lamprey"),
                         Species = factor(Species, 
                                          levels = c("Sockeye Salmon", "Coho Salmon", "Chum Salmon", 
                                                     "Coastal Cutthroat", "Dolly Varden", "Coastal Sculpin", 
                                                      "Pacific Lamprey"))) %>%   
                  group_by(Species) %>%
                          summarize(count = n(),
                                    meanFL  = mean(Length, na.rm = TRUE),
                                    minFL   = min(Length, na.rm = TRUE),
                                    maxFL   = max(Length, na.rm = TRUE),
                                    nWT     = sum(Weight > 0, na.rm=TRUE),
                                    meanWT  = mean(Weight, na.rm = TRUE),
                                    minWT   = min(Weight, na.rm = TRUE),
                                    maxWT   = max(Weight, na.rm = TRUE))

## Prepare Kable Table
kable(tally.table,
      "html",
      booktabs = T,
      align = c("c"),
      col.names = c("Species", "n", "Mean", "Min", "Max", "n", "Mean", "Min", "Max"),
      caption = "**Table 1. Summary of total number and biological data collected of fish captured since April 15, 2024.**",
      digits = 1) %>%
  #column_spec(c(1, 3,10,17), bold = TRUE) %>%
  kable_styling("striped", full_width =  FALSE,
                position = "center", 
                font_size = 12,
                fixed_thead = TRUE) %>%
  add_header_above(c(" " = 1, "Fork Length" = 4, "Weight (g)" = 4)) 
```


## Sockeye Salmon

The length-frequency distribution for Sockeye Salmon smolts is bimodal, indicating two groups of different sized fish are emigrating simultaneously. Scale aging will be required to assess what is contributing to this observed size difference. However, it is likely best explained by a difference in age class between modes (e.g., Age-1 and Age-2 smolts), though it is also possible that some of the difference is explained by a difference in productivity between rearing environments (e.g., some smolts may be the same age, but larger smolts reared in more productive environments than smaller smolts).

```{r smolt proportion, echo = FALSE, warning = FALSE, message = FALSE}
SK_AgeClass <- catch_dat %>% 
              filter(Species == "SK") %>%
              mutate(AgeClass =ifelse(Length>=120,"Age-2","Age-1"),
                     TotalCount = n())

SK_prop <- SK_AgeClass %>%
              group_by(AgeClass,TotalCount) %>%
              summarize(Count  = n(),
                        MeanFL = mean(Length, na.rm = TRUE),
                        SEFL   = sd(Length, na.rm = TRUE)/sqrt(Count),
                        MinFL  = min(Length, na.rm = TRUE),
                        MaxFL  = max(Length, na.rm = TRUE),
                        MeanWT = mean(Weight, na.rm = TRUE),
                        SEWT   = sd(Weight, na.rm = TRUE)/sqrt(Count),
                        MinWT  = min(Weight, na.rm = TRUE),
                        MaxWT  = max(Weight, na.rm = TRUE))%>%
              mutate(Prop = (Count/TotalCount)*100) %>% 
              select("AgeClass" ,"Count","Prop","MeanFL", "SEFL", "MinFL", "MaxFL", "MeanWT","SEWT", "MinWT","MaxWT")


kable(SK_prop,
      "html",
      booktabs = T,
      align = c("c"),
      col.names = c("Age Class", "n", "Prop (%)","Mean", "SE","Min", "Max", "Mean", "SE","Min", "Max"),
      caption = "**Table 2. Size of Age-1 and Age-2 Sockeye Salmon captured to date.**",
      digits = 1) %>%
  #column_spec(c(1, 3,10,17), bold = TRUE) %>%
  kable_styling("striped", full_width =  FALSE,
                position = "center", 
                font_size = 12,
                fixed_thead = TRUE) %>%
  add_header_above(c(" " = 3, "Fork Length" = 4, "Weight (g)" = 4)) 

```
To date, it appears that the "Age-1" Sockeye Salmon smolts range in size from 78 mm to 118 mm while Age-2 smolts are greater than 120 mm.

```{r SK_FL, echo = FALSE, warning = FALSE, fig.cap = "Length-frequency distribution for Sockeye Salmon captured in the Fulmore River.", message = FALSE, out.width= "1500px"}
## Prepare Plots ----
## * Length Frequency Plots  ----  
## ** Prepare Data
    ## Total tally by fork length
    all_catches <- catch_dat %>% 
                      filter(Species == "SK") %>%
                          group_by(Length) %>%
                          summarize(Count = n()) %>%
                          mutate(Date = "All Fish") %>%
                          select("Date","Length","Count")
    
    ## Daily tally by fork length
    daily_catch <- catch_dat %>%
                      filter(Species == "SK") %>%
                          group_by(Date, Length) %>%
                          summarize(Count = n()) %>%
                          select("Date","Length","Count") %>%
                          mutate(Date = as.character(Date))
    
    ## Weekly tally by fork length
    weekly_catch <- catch_dat %>%
                      filter(Species == "SK") %>%
                      mutate(Date = isoweek(ymd(Date))) %>%  # convert date to weeknumber
                      group_by(Date, Length) %>%
                        summarize(Count = n()) %>%
                        select("Date","Length","Count") %>%
                        mutate(Date = as.character(Date))

    ## Merge Total and Daily Tally              
    total_daily <- rbind(daily_catch, all_catches)
    
    ## Merge Total and Weekly Tally
    total_weekly <- rbind(weekly_catch, all_catches)

        
## ** Prepare Daily Length Frequency Plots
    ## Faceted by Day
    daily_fl.plot <- ggplot(total_daily) + 
                   geom_col(aes(x = Length, y = Count)) + 
                    labs(x = "Fork Lenght (mm)", y = "Frequency (# Fish)") +
                    scale_x_continuous(breaks = seq(76, 166, by = 2))+
                    scale_y_continuous(breaks = seq(0, 50, 2))+
                    plottheme() +
                    theme(axis.text.x = element_text(angle = 45, vjust = 1.2, hjust = 1)) +
                    facet_grid(Date ~.)

    ## Faceted by Week
    weekly_fl.plot <- ggplot(total_weekly) + 
                    geom_col(aes(x = Length, y = Count)) + 
                      labs(x = "Fork Lenght (mm)", y = "Frequency (# Fish)") +
                      scale_x_continuous(breaks = seq(76, 166, by = 2))+
                      scale_y_continuous(breaks = seq(0, 50, 2))+
                      plottheme() +
                      facet_grid(Date ~.)
    
## ** Select Figures to Display
    
    daily_fl.plot
    
    # weekly_fl
```                  


```{r total catches, out.width = "1500px", fig.cap = "Total catches of Sockeye Salmon, Coho Salmon and Chum Salmon per day.", echo = FALSE, message = FALSE, warning = FALSE}
## * Total Captures Plot ----
    ## Prep Blank Date Data  
    date_seq <- data.frame(Date = rep(seq(ymd('2024-04-15'), ymd('2024-06-15'), by = 'days'), each = 3)) %>%
                mutate(Month = strftime(Date, format = "%m"),
                       Month = as.numeric(Month),
                       Week  = strftime(Date, format = "%V"),
                       Species = rep(c("SK","CO","CM"), length.out = 186))

    ## Prepare Data
    running_tally <- catch_dat %>% 
                       group_by(Species, Date) %>%
                          summarize(Count = n()) %>% 
                          filter(Species == "SK" | Species== "CO" | Species == "CM")

    
    daily_tally <- left_join(date_seq, running_tally, by = c("Date", "Species")) %>%
                   mutate(Count = coalesce(Count, 0),
                          Date  = as.Date(Date),
                          Species = recode(as.factor(Species), 
                                         SK        = "Sockeye Salmon",
                                         CO        = "Coho Salmon",
                                         CM        = "Chum Salmon"),
                                       # CT        = "Coastal Cutthroat",
                                       # 'DV-BT'   = "Dolly Varden",
                                       # CC        = "Coastal Sculpin",
                                       # Lamprey   = "Pacific Lamprey"),
                         Species = factor(Species, 
                                          levels = c("Sockeye Salmon", "Coho Salmon", "Chum Salmon")))
                                                      # "Coastal Cutthroat", "Dolly Varden", "Coastal Sculpin", 
                                                      # "Pacific Lamprey"))) 

    ## Prepare Plot
    daily_tally.plot <- ggplot(daily_tally) +
                          geom_col(aes(x = Date, y = Count)) +
                          labs(x = "Date", y = "Total Captures (# Fish)") +
                          scale_x_date(date_breaks = "1 week", date_labels = "%d-%b") +
                          plottheme() +
                          facet_grid(Species ~.)
    
    daily_tally.plot

```


